Automating Verification of MPC Security: $\metaprot$ to $\minifed$

### 2023-07-09 popl24/automation.tex
checkpoint

We can thus automatically verify security properties of examples considered
in previous sections. 
\begin{lemma}
  $\prog_{\ref{example-he}}$ is passive secure,
  $\prog_{\ref{example-OT}}$ is passive secure assuming client 3 is
  honest, and $\prog_{\ref{example-otp}}$ satisfies probabilistic
  noninterference and hence perfect secrecy as defined in
  \ref{barthe2019probabilistic}.
\end{lemma}
\begin{proof}
  We have certified $\nimo(prog_{\ref{example-he}})$, $\nimo(prog_{\ref{example-OT}})$ assuming
  $3 \in H$, and $\pni(\prog_{\ref{example-otp}})$. The result follows by Lemmas
  \ref{lemma-pni} and \ref{lemma-ni}.
\end{proof}

@@

While the brute force method is feasible for smaller programs, it is
not feasible for larger programs such as large YGC or GMW circuits
since the size of $\runs(\prog)$ is exponential in the size of the
random tape plus the number of input secrets.

However, circuit abstractions are based on component gates which
define subprotocols that we can consider in isolation. Out approach to
verifying circuits strategy is thus to show that any circuit built
from independently certified gates satisfies $\NIMO$. Intuitively, we
can consider any gate as a mini-program and certify intensional
properties of its isolated program distribution. These intensional
properties are guaranteed secure in circuit compositions, as we
demonstrate using methods based on probabilistic separation logic
\ref{barthe2019probabilistic}. While this last step is manual, our
gate certifications are defined generally and can be used to certify
new gate implementations, and verification applies to a circuit of any
size. In Section \ref{section-composition} we will demonstrate our
approach using YGC and GMW as examples.

@@

Since we certify properties of gates in isolation with witness inputs,
our metatheory is concerned with preservation of important invariants
when certified components are ``spliced'' into larger program
contexts.  One important reasoning principle is related to
noninterference properties- specifically, if we show that gate views
enjoy a noninterference property wrt their free (input) variables,
probabilistic independence propagates to other variables in the
preceding program.
\begin{lemma}[Noninterference]
  \label{lemma-noninterference}
  Given $\prog_1;\prog_2$ and $X = \iov(\prog_2) - \vdefs(\prog_2)$ and
  $Y \subseteq \vdefs(\prog_2)$. If $\sep{\progd(\prog_1;\prog_2)}{X}{Y}$
  then $\sep{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}{Y}$.
\end{lemma}

Also, when we certify gates in isolation with witness inputs, we can
``splice'' these properties into larger program contexts using the
following substitution Lemmas, showing that the substitution of
wire inputs for witness inputs preserves both separation and correlation.
\begin{lemma}[Substitution$*$]
  \label{lemma-substitution}
  If $\sep{\progtt(\prog_2)}{\{ f \}}{X}$ and
  $\prog_1;\prog_2[\be/f]$ is safe with $\vars(\prog_1,\be) \cap
  X = \varnothing$ then
  $\sep{\progtt(\prog_1;\prog_2[\be/f])}{\vars(\be)}{X}$.
\end{lemma}

@@

We note that while the Frame rule of \cite{barthe2019probabilistic}
supports local reasoning in larger program contexts, it does not
support this splicing method, and our Lemmas \ref{lemma-substitution}
and \ref{lemma-substitution-sim} are both novel.

>>>>

We can thus automatically verify security properties of examples considered
in previous sections, observing that passive security of $\prog_{\ref{example-OT}}$
has not been previously demonstrated.
\begin{lemma}
  $\prog_{\ref{example-he}}$, $\prog_{\ref{example-gmw-andcircuit}}$,
  $\prog_{\ref{example-ygc-andcircuit}}$ are all passive secure,
  $\prog_{\ref{example-OT}}$ is passive secure assuming client 3 is
  honest, and $\prog_{\ref{example-otp}}$ satisfies probabilistic
  noninterference and hence perfect secrecy as defined in
  \ref{barthe2019probabilistic}.
\end{lemma}
\begin{proof}
  We have certified $\nimo(prog_{\ref{example-he}})$, $\nimo(\prog_{\ref{example-gmw-andcircuit}})$,
  $\nimo(\prog_{\ref{example-ygc-andcircuit}})$, and $\nimo(prog_{\ref{example-OT}})$ assuming
  $3 \in H$, and $\pni(\prog_{\ref{example-otp}})$. The result follows by Lemmas
  \ref{lemma-pni} and \ref{lemma-nimo}.
\end{proof}

@@

While the brute force method is feasible for smaller programs,
including small circuits such as Examples
$\prog_{\ref{example-gmw-andcircuit}}$ and
$\prog_{\ref{example-ygc-andcircuit}}$, it is not for larger programs
such as YGC or GMW circuits since the size of $\runs(\prog)$ is
exponential in the size of the random tape plus the number of input
secrets, and circuits can be quite large in practice
\cite{kreuter2012billion}.  However, the YGC and GMW protocols are
examples of a common idiom in MPC protocol design- arbitrarily large
circuits are built up from smaller boolean or arithmetic gates that
operate on encodings of input secrets.

Our approach is thus to design gate certifications that demonstrably
guarantee security for circuits built from them. Intuitively, we can
consider any gate as a mini-program and certify intensional properties
of its isolated program distribution. These intensional properties are
guaranteed secure in circuit compositions, as we demonstrate using
methods based on probabilistic separation logic
\ref{barthe2019probabilistic}. While this last step is manual, our
gate certifications are defined generally and can be used to certify
new gate implementations, and verification applies to a circuit of any
size. In Section \ref{section-composition} we will demonstrate our
approach using YGC and GMW as examples.

@@

Since we certify properties of gates in isolation with witness inputs,
our metatheory is concerned with preservation of important invariants
when certified components are ``spliced'' into larger program
contexts.  One important reasoning principle is related to
noninterference properties- specifically, if we show that gate views
enjoy a noninterference property wrt their free (input) variables,
probabilistic independence propagates to other variables in the
preceding program.
\begin{restatable}[Noninterference]{lemma}{noninterference}
  \label{lemma-noninterference}
  Given $\prog_1;\prog_2$ and $X = \iov(\prog_2) - \vdefs(\prog_2)$ and
  $Y \subseteq \vdefs(\prog_2)$. If $\sep{\progd(\prog_1;\prog_2)}{X}{Y}$
  then $\sep{\progd(\prog_1;\prog_2)}{\iov(\prog_1)}{Y}$.
\end{restatable}

Also, when we certify gates in isolation with witness inputs, we can
``splice'' these properties into larger program contexts using the
following substitution Lemmas, showing that the substitution of
wire inputs for witness inputs preserves both separation and correlation.
\begin{restatable}[Substitution$*$]{lemma}{substar}
  \label{lemma-substitution}
  If $\sep{\progtt(\prog_2)}{\{ f \}}{X}$ and
  $\prog_1;\prog_2[\be/f]$ is safe with $\vars(\prog_1,\be) \cap
  X = \varnothing$ then
  $\sep{\progtt(\prog_1;\prog_2[\be/f])}{\vars(\be)}{X}$.
\end{restatable}

@@

We note that while the Frame rule of \cite{barthe2019probabilistic}
supports local reasoning in larger program contexts, it does not
support this splicing method, and our Lemmas \ref{lemma-substitution}
and \ref{lemma-substitution-sim} are both novel. Supplementary proofs
are demonstrated in Appendix \ref{section-proofs}.

### 2023-07-09 popl24/bruteforce.tex
checkpoint

\section{Brute Force Verification}
\label{section-bruteforce}

We have shown how our formulation of program distributions, a precise
definition of intensional protocol behavior, can be used to enforce
the extensional property of MPC security. This intensional description
also has the benefit of being flexible- we can use the same mechanisms
to describe $\NIMO$, noninterference, perfect secrecy, violations of
obliviousness, etc. These mechanisms can also be exploited to support
automated program analysis.

We initially explore a brute force approach to automatically verifying
program behaviour, the first step of which is to compute the basic
distribution of a program $\prog$. The program distribution
$\progd(\prog)$ can obviously be automatically derived from that, as
can be any of its marginal or conditional distributions. In this
Section, we consider two automated mechanisms for deriving
$\progd(\prog)$ in this way. The first is through a straightforward
computation of all final program memories.  The second approach
rewrites $\minifed$ programs to stratified Datalog programs, that are
amenable HPC optimizations such as parallelization and GPU matrix
computations as shown in recent work
\cite{sakama2017linear,aspis2018linear,nguyen2022enhancing,nguyen2021efficient}.

The basic distribution of a program $\prog$ is based on the
final memories resulting from any run of $\prog$, denoted
$\runs(\prog)$.
\begin{definition}
  Given $\prog$ with $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$:
  $$
  \runs(\prog) \defeq \{ \store \in \mems(S\cup F \cup V) \mid \config{\store_{S \cup F}}{\prog} \redxs \config{\store}{\varnothing} \}
  $$
\end{definition}
As we've observed, $|\runs(\prog)| = 2^{|S \cup F|}$, so the size of
$\runs(\prog)$ is exponential in the amount of randomness used in the
protocol which generally can be used as a proxy for the protocol
size. In Section \ref{section-composition} we will show how the
methods described here can be used to individually verify
compositional properties of small program components, that guarantee
security properties in larger protocols that combine them. Thus, while
scalability to program size is a clear challenge for our brute force
techniques alone, our ultimate goal is to use them for verification of
compositional properties in larger programs.

@@

In order to automatically obtain $\runs(\prog)$, we can do a left
folding of $\solve$ across $\prog$, inductively extending memories
with valid view assignments in the order of their definition.
We can then use this to compute $\progd(\prog)$. We denote
these computations as $\cruns$ and $\cprogd$ respectively.
\begin{lemma}
  \label{lemma-cruns}
  Given $\prog$ where $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$. Define:
  \begin{eqnarray*}
    {tt}\ \ \stores\ (\eassign{\view{\cid}{w}}{\be}) &\defeq& \begin{array}{l}
      \mathrm{let}\ \stores' = \solve{\stores}{\be} \ \mathrm{in}\\
      \ \ \{\extend{\store}{\view{\cid}{w}}{1} \mid \store \in \stores' \}\ \cup\\
      \ \ \{\extend{\store}{\view{\cid}{w}}{0} \mid \store \in \stores - \stores' \}\end{array}\\[2mm]
    \cruns(\prog) &\defeq& \mathit{foldl}\ {tt}\ \mems(S \cup F)\ \prog
  \end{eqnarray*}
  Then $\cruns(\prog) = \runs(\prog)$.
\end{lemma}
\begin{proof}
  By Lemma \ref{lemma-solves} and induction on the length of $\prog$. 
\end{proof}
\begin{lemma}
  \label{lemma-cprogd}
  Given $\prog$ where $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$. Define
  $\cprogd(\prog)$ as the distribution $\margd{\pdf{S \cup F \cup V}}{S\cup V}$,
  where:
  $$
  \pdf{S \cup F \cup V}(\store) = \begin{cases}1/2^{|S \cup F|} & \text{if\ } \store \in
    \cruns(\prog) \\ 0 & \text{otherwise} \end{cases} 
  $$
  Then:
  $$
  \cprogd(\prog) = \progd(\prog)
  $$
\end{lemma}
\begin{proof}
  Immediately by Lemma \ref{lemma-cruns} and Definition \ref{def-progd}.
\end{proof}

We have used the code for $\cruns$ and $\cprogd$ extensively to design and
verify the examples in this paper. We will discuss other examples in
Section \ref{section-composition}, but here we note that we've verified
passive security of Examples \ref{example-he} and \ref{example-OT}-
for the former, this is well-known so the result is a simple proof-of-concept,
but the latter is a new (though unsurprising) result, having not been
directly proven in \cite{barthe2019probabilistic}. The method for
verifying $\NIMO$ described here generalizes to any $\prog$. 
\begin{lemma}[Automated Verification of Passive Security]
  \label{lemma-bruteforce-nimo}
  $\prog_{\ref{example-he}}$ and $\prog_{\ref{example-OT}};\eassign{\outv}{\vx{2}{mc}}$ are
  passive secure.
\end{lemma}
\begin{proof}
  For both $\prog \in \{
  \prog_{\ref{example-he}},\prog_{\ref{example-OT}};\eassign{\outv}{\vx{2}{mc}}\}$
  we compute $\cruns(\prog)$ and $\cprogd(\prog)$ and then verify
  $\NIMO(\prog)$ using the following algorithm. Given $\iov(\prog) = S
  \cup V$ and $\flips(\prog) = F$, we enumerate all memories $\store
  \in \cruns(\prog)$, then enumerate all partitions of clients into
  sets $H$,$C$ such that $|C| \le |H|$ and $0 \in C$, and then check:
  $$
  \cprogd(\prog)(\store_{S_H}|\store_{S_C \cup \{\outv\}}) =
  \cprogd(\prog)(\store_{S_H}|\store_{S_C \cup V_C})
  $$
  for each $H,C,\store$. The result follows by Lemma \ref{lemma-cprogd}, Lemma \ref{lemma-nimo}, and
  Theorem \ref{theorem-nimo}.
\end{proof}

>>>>

\section{Brute Force Computation of Program Distributions}
\label{section-bruteforce}

We first explore a brute force approach to automatically verifying
program behaviour, the first step of which is to compute the basic
distribution of a program $\prog$. The program distribution
$\progd(\prog)$ can obviously be automatically derived from that, as
can be any of its marginal or conditional distributions. In this
Section, we consider two automated mechanisms for deriving
$\progd(\prog)$ in this way. The first is through a straightforward
computation of all final program memories.  The second approach
rewrites $\minifed$ programs to stratified Datalog programs, that are
amenable HPC optimizations such as parallelization and GPU matrix
computations as shown in recent work
\cite{sakama2017linear,aspis2018linear,nguyen2022enhancing,nguyen2021efficient}.

%The basic distribution of a program $\prog$ is based on the
%final memories resulting from any run of $\prog$, denoted
%$\runs(\prog)$.
%\begin{definition}
%  Given $\prog$ with $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$:
%  $$
%  \runs(\prog) \defeq \{ \store \in \mems(S\cup F \cup V) \mid \config{\store_{S \cup F}}{\prog} \redxs \config{\store}{\varnothing} \}
%  $$
%\end{definition}

As we've observed, $|\runs(\prog)| = 2^{|S \cup F|}$, so the size of
$\runs(\prog)$ is exponential in the amount of randomness used in the
protocol which generally can be used as a proxy for the protocol
size. In Section \ref{section-composition} we will show how the
methods described here can be used to individually verify
compositional properties of small program components, that guarantee
security properties in larger protocols that combine them. Thus, while
scalability to program size is a clear challenge for our brute force
techniques alone, our ultimate goal is to use them for verification of
compositional properties in larger programs.

@@

In order to automatically obtain $\runs(\prog)$, we can do a left
folding of $\solve$ across $\prog$, inductively extending memories
with valid view assignments in the order of their definition.
We can then use this to compute $\progd(\prog)$. We denote
these computations as $\cruns$ and $\cprogd$ respectively.
\begin{lemma}
  \label{lemma-cruns}
  Given $\prog$ where $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$. Define:
  \begin{eqnarray*}
    {tt}\ \ \stores\ (\eassign{\view{\cid}{w}}{\be}) &\defeq& \begin{array}{l}
      \mathrm{let}\ \stores' = \solve{\stores}{\be} \ \mathrm{in}\\
      \ \ \{\extend{\store}{\view{\cid}{w}}{1} \mid \store \in \stores' \}\ \cup\\
      \ \ \{\extend{\store}{\view{\cid}{w}}{0} \mid \store \in \stores - \stores' \}\end{array}\\[2mm]
    \cruns(\prog) &\defeq& \mathit{foldl}\ {tt}\ \mems(S \cup F)\ \prog
  \end{eqnarray*}
  Then $\cruns(\prog) = \runs(\prog)$.
\end{lemma}
\begin{proof}
  By Lemma \ref{lemma-solves} and induction on the length of $\prog$. 
\end{proof}
\begin{lemma}
  \label{lemma-cprogd}
  Given $\prog$ where $\iov(\prog) = S \cup V$ and $\flips(\prog) = F$. Define
  $\cprogd(\prog)$ as the distribution $\pmf$
  where for all $\store \in \mems(S\cup F \cup V)$:
  $$
  \pmf(\store) = \begin{cases}1/2^{|S \cup F|} & \text{if\ } \store \in
    \cruns(\prog) \\ 0 & \text{otherwise} \end{cases} 
  $$
  Then:
  $$
  \cprogd(\prog) = \progd(\prog)
  $$
\end{lemma}
\begin{proof}
  Immediately by Lemma \ref{lemma-cruns} and Definition \ref{def-progd}.
\end{proof}

%%% I moved the following to automation.tex - chris
\begin{comment}
We have used the code for $\cruns$ and $\cprogd$ extensively to design and
verify the examples in this paper. We will discuss other examples in
Section \ref{section-composition}, but here we note that we've verified
passive security of Examples \ref{example-he} and \ref{example-OT}-
for the former, this is well-known so the result is a simple proof-of-concept,
but the latter is a new (though unsurprising) result, having not been
directly proven in \cite{barthe2019probabilistic}. The method for
verifying $\NIMO$ described here generalizes to any $\prog$. 
\begin{lemma}[Automated Verification of Passive Security]
  \label{lemma-bruteforce-nimo}
  $\prog_{\ref{example-he}}$ and $\prog_{\ref{example-OT}};\eassign{\outv}{\vx{2}{mc}}$ are
  passive secure.
\end{lemma}
\begin{proof}
  For both $\prog \in \{
  \prog_{\ref{example-he}},\prog_{\ref{example-OT}};\eassign{\outv}{\vx{2}{mc}}\}$
  we compute $\cruns(\prog)$ and $\cprogd(\prog)$ and then verify
  $\NIMO(\prog)$ using the following algorithm. Given $\iov(\prog) = S
  \cup V$ and $\flips(\prog) = F$, we enumerate all memories $\store
  \in \cruns(\prog)$, then enumerate all partitions of clients into
  sets $H$,$C$ such that $|C| \le |H|$ and $0 \in C$, and then check:
  $$
  \cprogd(\prog)(\store_{S_H}|\store_{S_C \cup \{\outv\}}) =
  \cprogd(\prog)(\store_{S_H}|\store_{S_C \cup V_C})
  $$
  for each $H,C,\store$. The result follows by Lemma \ref{lemma-cprogd}, Lemma \ref{lemma-nimo}, and
  Theorem \ref{theorem-nimo}.
\end{proof}

\end{comment}

### 2023-07-09 popl24/hyperprop.tex
checkpoint

Probabilistic noninterference can also be restated to formally capture a
basic intuition that noninterfering programs do not provide
the adversary with any clues about honest secrets through corrupt
views, as in the following Lemma. This is similar to asserting that
adversarial knowledge is not increased by the execution of a program as in formulations
of so-called gradual release \cite{4223226}. 
\begin{lemma}
  \label{lemma-pni)
  Given a program $\prog$ with $\iov(\prog) = (S,V)$, 
  $\prog$ satisfies probabilistic noninterference iff for all
  $\store \in \mems(S \cup V)$:
  $$\condd{\progd(\prog)}{S_H}{\store_{S_C}} =
    \condd{\progd(\prog)}{S_H}{\store_{(S\cup V)_C}} $$
\end{lemma}

>>>>

Probabilistic noninterference can also be restated to formally capture a
basic intuition that noninterfering programs do not provide
the adversary with any clues about honest secrets through corrupt
views, as in the following Lemma. This is similar to asserting that
adversarial knowledge is not increased by the execution of a program as in formulations
of so-called gradual release \cite{4223226}. 
\begin{lemma}
  \label{lemma-pni}
  Given a program $\prog$ with $\iov(\prog) = (S,V)$, 
  $\prog$ satisfies probabilistic noninterference iff for all
  $\store \in \mems(S \cup V)$:
  $$\condd{\progd(\prog)}{S_H}{\store_{S_C}} =
    \condd{\progd(\prog)}{S_H}{\store_{(S\cup V)_C}} $$
\end{lemma}

### 2023-07-09 popl24/intro.tex
checkpoint

Secure Multi-Party Computation (MPC) protocols support data privacy in
important modern, distributed applications such as privacy-preserving
machine learning \cite{li2021privacy, knott2021crypten, koch2020privacy, liu2020privacy} and Zero-Knowledge proofs in blockchains \cite{ishai2009zero, lu2019honeybadgermpc, gao2022symmeproof, tomaz2020preserving}
\cnote{Need some banger cites here, maybe more MPC hype.}. MPC methods
have been developed by the cryptography community for many years,
while receiving the attention of the programming languages community
only relatively recently. In contrast, information flow security has
received significant attention in PL theory and practice especially
since the turn of the century \cite{1159651}, including a menagerie of
variants, enforcement mechanisms, and programming frameworks. Much of
this has been enabled by the unified metatheory of
\emph{hyperproperties} \cite{10.5555/1891823.1891830} that establishes
a common conceptual framework for reasoning about and implementing
systems with information flow security.

@@

In Section \ref{section-minicat} we develop a new probabilistic programming
language $\minifed$ for defining synchronous distributed protocols
over the binary field. The syntax and semantics provides a succinct
account of \emph{views} and (synchronous) messaging between protocol
\emph{clients}. In this Section, we consider small example protocols
including oblivious transfer, communications encryption with one-time
pads, and additive secret sharing. In Section \ref{section-pmf} we
formulate \emph{program distributions}, a formalism for expressing
dependencies between secrets and views, useful for expressing
intensional characteristics of protocols and precisely quantifying
allowable information leakage in terms of protocol outputs. On this
basis we then formulate a novel hyperproperty of program execution
traces, called \emph{noninterference modulo output ($\NIMO$)} that we
demonstrate implies passive simulator security. In Section
\ref{section-bruteforce} we define a brute force mechanism for
verifying passive security in $\minifed$ protocols, that is also
amenable to HPC optimizations via translation of $\minifed$ protocols
into Datalog \cnote{I think this Section can be on the chopping block
  for space savings, since it's based on truth table
  computation/Datalog semantics.}.

>>>>

Secure Multi-Party Computation (MPC) protocols support data privacy in
important modern, distributed applications such as privacy-preserving
machine learning \cite{li2021privacy, knott2021crypten, koch2020privacy, liu2020privacy} and Zero-Knowledge proofs in blockchains \cite{ishai2009zero, lu2019honeybadgermpc, gao2022symmeproof, tomaz2020preserving}. MPC methods
have been developed by the cryptography community for many years,
while receiving the attention of the programming languages community
only relatively recently. In contrast, information flow security has
received significant attention in PL theory and practice especially
since the turn of the century \cite{1159651}, including a menagerie of
variants, enforcement mechanisms, and programming frameworks. Much of
this has been enabled by the unified metatheory of
\emph{hyperproperties} \cite{10.5555/1891823.1891830} that establishes
a common conceptual framework for reasoning about and implementing
systems with information flow security.

@@

In Section \ref{section-minicat} we develop a new probabilistic programming
language $\minifed$ for defining synchronous distributed protocols
over the binary field. The syntax and semantics provides a succinct
account of \emph{views} and (synchronous) messaging between protocol
\emph{clients}. In this Section, we consider small example protocols
including oblivious transfer, communications encryption with one-time
pads, and additive secret sharing. In Section \ref{section-pmf} we
formulate \emph{program distributions}, a formalism for expressing
dependencies between secrets and views, useful for expressing
intensional characteristics of protocols and precisely quantifying
allowable information leakage in terms of protocol outputs. On this
basis we then formulate a novel hyperproperty of program execution
traces, called \emph{noninterference modulo output ($\NIMO$)} that we
demonstrate implies passive simulator security. In Section
\ref{section-bruteforce} we define a brute force mechanism for
verifying passive security in $\minifed$ protocols, that is also
amenable to HPC optimizations via translation of $\minifed$ protocols
into Datalog.

### 2023-07-09 popl24/metalang.tex
checkpoint

\section{A Protocol Metalanguage}
\label{section-metalang}

@@

>>>>

\section{The $\metaprot$ Metalanguage}
\label{section-metalang}

@@

\subsection{Example: GMW}
\label{section-metalang-gmw}

\input{gmw}

\begin{fpfig}[t]{GMW library type annotations.}{fig-gmw-types}
{\footnotesize
  \begin{verbatimtab}
   encodegmw   : string(gid) * string(gid)

   andtablegmw : { k = bool[i]; p = bool[i] }

   andgmw      : string(gid) *  { c1 = bool[1]; c2 = bool[2] } * { c1 = bool[1]; c2 = bool[2] }

   decodegmw   : { c1 = bool[1]; c2 = bool[2] }  \end{verbatimtab}
}
\end{fpfig}

The GMW protocol uses secret sharing to represent data flowing through
circuits. In the 2-party case, clients 1 and 2 each share their input
secrets, and use those shares to represent inputs to gates. Outputs
are also represented as shares. We refer to the pair of shares
representing any particular value as a \emph{wire value}, and
we represent them via records of the form
$
\ttt{\{c1 = } v_1\ttt{;c2 = } v_2\ttt{\}} 
$
where $v_1$ and $v_2$ are client 1 and 2's shares respectively.

For full details of the GMW protocol the reader is referred to
\cite{evans2018pragmatic}. Our implementation libary is shown in
Figure \ref{fig-gmw}, with type signatures for the library functions
shown in Figure \ref{fig-gmw-types}. We show the And gate since it is
an interesting component. The Figure includes the
following top-level functions:
\begin{itemize}
\item \ttt{encodegmw}: This function encodes client 1's and client 2's
  secret bits called $\ttt{s[1,s1]}$ and $\ttt{s[1,s2]}$ into two
  distinct wire values (pairs of shares).
\item \ttt{andgmw}: This function defines the gate $\ttt{g}$, the
  identifier $\ttt{g}$ being used to distinguish randomness used
  within.  In our version client 1 builds the output table (using
  \ttt{andtablegmw})and transfers the correct output share to client 2
  using 1-out-of-4 OT as per standard GMW protocol.
\item \ttt{decode}: This function decodes and publishes a wire value
  by $\ttt{xor}$ing the shares. Note that this requires both client 1
  and 2 to publicize their shares.
\end{itemize}
\begin{example}
  \label{example-gmw-andcircuit}
The following program uses our GMW library to define
a circuit with a single And gate and input secrets $\ttt{s1}$ and
$\ttt{s2}$ from client's 1 and 2 respectively. 
\begin{verbatimtab}
  let ss = encodegmw(s1,s2) in
  v[0,output] := decode(andgmw(0,ss.shares1,ss.shares2))
\end{verbatimtab}
\end{example}

\subsection{Example: YGC}
\label{section-metalang-ygc}

\input{ygc}

In Figures \ref{fig-ygc-aux} and \ref{fig-ygc-gates} we define a
codebase for Yao's garbled circuits (YGC). This definition follows the
\emph{point-and-permute} method described in \cite{evans2018pragmatic}
and elsewhere, to which the reader is referred for more in-depth discussion.
In this implementation client 2 is the \emph{garbler} and
client 1 is the \emph{evaluator}. The garbler builds the garbled
tables and shares them with the evaluator, who then evaluates
the gate in an oblivious fashion until the final public output is
generated through decryption. This definition is well-typed,
with input type annotations for top-level functions listed in
Figure \ref{fig-ygc-types}. Well-typed programs using these
libraries are therefore guaranteed to yield safe $\minifed$
programs.

\emph{Wire labels} are fundamental to YGC, and essentially represent
gate output values in an encrypted form. In our definition, wire
labels are represented by records $\ttt{\{ k = }\beta_1\ttt{; p =
}\beta_2\ttt{ \}}$, where $\ttt{k}$ is the \emph{key bit} and
$\ttt{p}$ is the \emph{pointer bit}, and $\beta_1$ and $\beta_2$ are
flips. Flips in each output wire label are owned by the garbler and
are unique per gate by definition of their identifying string, and the
representation of $0$ is the negation of $1$. For example, here is the
representation of 1 and 0 respectively in the output wire label for a
hypothetical gate 6:
\begin{mathpar}
  \ttt{\{ k = flip[2,gate:6.k]; p =  flip[2,gate:6.p]] \}}

  \ttt{\{ k = not flip[2,gate:6.k]; p =  not flip[2,gate:6.p]] \}}
\end{mathpar}
The pointer bits in wire labels are used to select permuted rows in
table garblings. The key bits are used to identify a unique key for
table row in each garbled gate. Intuitively, if $\beta_1$ and
$\beta_2$ are either key or pointer bits encoding 1 on two input wire
labels to a binary gate, rows and keys in the gate are enumerated in
the order:
$$
\neg\beta_1\neg\beta_2,\ \neg\beta_1\beta_2,\ \beta_1\neg\beta_2,\ \beta_1\beta_2
$$

In our implementation, gates are wired together using gate
identifiers, which are strings $w$. Top-level functionality in Figures
\ref{fig-ygc-aux} and \ref{fig-ygc-gates} includes the following:
\begin{itemize}
\item \ttt{andgate}: This defines a subprotocol for the garbler
  to define a garbled gate $\ttt{gid}$ with input wires from gates
  $\ttt{ga}$ and $\ttt{gb}$. The garbler generates keys and garbles
  the rows in YGC fashion, them with client $1$ in
  views in a standard form. For example, the view for
  a hypothetical gate 6, row 2 garbled truth table is $\ttt{v[1,gate:6tt2]}$.
  We note that garbled gates of other binary operators can be obtained with
  replacement of $\ttt{andtable}$ with appropriate garbled table definitions. 
\item \ttt{evalgate}: This defines a subprotocol for the evaluator to
  evaluate gate $\ttt{gid}$ given input wire values $\ttt{wva}$ and
  $\ttt{wvb}$.
\item \ttt{garbledecode} and \ttt{evaldecode}: The former function
  defines the garbler's protocol for encrypting the circuit
  output from final gate $\ttt{gid}$, and the latter defines
  the evaluator's output decryption protocol.
\item \ttt{encode}: This defines the initial phase of the protocol,
  where the evaluator receives the wire value from their own
  secret $\sx{1}{sa}$ via $\ttt{OT}$, and the garbler communicates
  the wire value for their own secret $\sx{2}{sb}$ directly.
\end{itemize}
\begin{example}
  \label{example-ygc-andcircuit}
The following program uses our YGC library to define
a circuit with a single and gate and input secrets $\ttt{s1}$ and
$\ttt{s2}$ from client's 1 and 2 respectively. 
\begin{verbatimtab}
  andgate(0,s1,s2);
  garbledecode(0);
  let secrets = encode(s1,s2) in
  v[0,output] := decode(evalgate(0, secrets.wv1, secrets.wv2))
\end{verbatimtab}
\end{example}
